Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2019-04-03 thesis/chapters/ch_chapter4.tex
finalized draft of chapter 4 with references

\chapter[Mixing Properties of Individual Enrichment Sources as a Function of Energy]{Mixing Properties of Individual Enrichment Sources as a Function of Energy\label{ch:chapter4}}
\let\thefootnote\relax\footnotetext{This section contains unpublished work-in-progress}

@@

The elemental abundances of a galaxy over time are sensitive to the nuclear physics and stellar astrophysics that determines which stars make what elements and when. The stellar abundance patters within a given galaxy also depend on the details of how those metals are released into the ISM through various forms of stellar feedback, and the hydrodynamic interactions that ultimately mix those elements into the ISM or eject them from the galaxy in galactic winds. Stellar abundance patterns in old, metal poor systems are the result of enrichment from substantially fewer sources than younger, more metal rich stellar populations. Because of this, studying abundance patterns in these environments allows us to place constraints on the yields of individual enrichment events and the nucleosynthetic sites of each element. The lowest mass dwarf galaxies in the Local Group, UFDs, offer some of the best constraints on these processes. These environments can potentially be used to trace and constrain enrichment from the first stars \citep[e.g.][]{FrebelBromm2012,Ji2015,Ritter2015,Jeon2017,Hartwig2018}. and have been used to place constraints on the variety of possible astrophysical sources of r-process enrichment \citep[e.g.][]{Ji2016a,Ji2016b,Ji2018a,Ji2019,Tsujimoto2017,Ojima2018,Duggan2018}. It is likely that chemical enrichment in more massive, metal rich systems with ongoing star formation like the Milky Way is much more uniform, averaging over the variety of complex hydrodynamics and feedback-driven turbulence interactions that may be driving much of the significant variance in stellar abundances seen at low metallicity. Stochasticity, which has been shown to be important for setting the width of stellar abundance patterns low metallicity environments of the Milky Way \citep[e.g.][]{Cescutti2008,Cescutti2014}, plays an even greater role in the evolution of these low mass dwarf galaxies \citep{Applebaum2018}. This complicates the interpretation of abundance patters in these galaxies, particularly for constraining individual enrichment events.

Recent cosmological zoom simulations have attained sufficient resolution to follow the evolution of individual UFDs (or their progenitors at high redshift) (references) and begin investigating their chemical properties \citep{Jeon2017,Corlies2018,Escala2018}. Additional works have conducted more direct investigations into what drives the enrichment process for individual sources using both hydrodynamics simulations \citep{PanScannapiecoScalo2013,Ritter2015,Safarzadeh2017,Hirai2015,Hirai2017,Emerick2018b,HaynesKobayashi2019} and semi-analytic models \citep{Beniamini2018,KrumholzTing2018}. In \cite{Emerick2018b} we demonstrate that elements released through AGB winds have both wider abundance distributions in the ISM and lower ejection fractions from the galaxy than elements released in supernovae. It is becoming increasingly clear that enrichment is quite inhomogeneous in low mass dwarf galaxies (refs?), and also that the behavior of metals from any one enrichment event depends on a significant number of complex interactions. In spite of this progress, there is still substantial work to be done in understanding the physical processes that drive both evolution of both the mean and width of stellar abundances in low-mass dwarf galaxies.

@@

However, the metals from each source do gradually trend towards a fraction of $\sim$0.8 by the end of the 150~Myr simulation time, corresponding to the mass fraction of the CNM at this point in the galaxy's evolution (see Figure~\ref{ch1:figISM_evolution}). This trend -- that the fraction of metals contained in a given phases tends towards the mass fraction of that phase -- is true across all phases in the simulation. Therefore, we can consider that the metals for each source are well-mixed across the \textit{phases} of the ISM on timescales of $\sim$100-150~Myr. \textbf{Compute the cooling timescale of each phase (maybe with a phase diagram?) and also the dynamical timescale and make comparisons here.}. These two results suggest that the delay-time between when metals are available to enrich ongoing star formation is on order of $\sim$10~Myr across sources, and only significant for the lowest (E46, AGB-like) energy sources. Although well-mixed across phases, we emphasize that this does not imply that the abundances are the same across phases nor that the metals are spatially well-mixed across the galaxy. We investigate this second point further below.

@@

What this does suggest, however, is that in utilizing stars with unusual abundances at low metallicities to infer the yields of exotic enrichment sources, the spread of abundances in those stars may provide an important insight into the origin of those metals. For example, r-process enrichment from exotic enrichment events like hypernovae should manifest itself as more well-mixed than if the elements originated in lower energy NS-NS mergers.
\textbf{Go back to reticulum papers and see what the spreads on these are}

@@

\textbf{Discuss Mean-median evolution here once AGB400 is up}

>>>>

\chapter[Mixing Properties of Individual Enrichment Sources]{Mixing Properties of Individual Enrichment Sources\label{ch:chapter4}}
\begingroup
\let\thefootnote\relax\footnotetext{This section contains unpublished work-in-progress}
\endgroup

@@

The elemental abundances of a galaxy over time are sensitive to the nuclear physics and stellar astrophysics that determines which stars make what elements and when. The stellar abundance patters within a given galaxy also depend on the details of how those metals are released into the ISM through various forms of stellar feedback, and the hydrodynamic interactions that ultimately mix those elements into the ISM or eject them from the galaxy in galactic winds. Stellar abundance patterns in old, metal poor systems are the result of enrichment from substantially fewer sources than younger, more metal rich stellar populations. Because of this, studying abundance patterns in these environments allows us to place constraints on the yields of individual enrichment events and the nucleosynthetic sites of each element. The lowest mass dwarf galaxies in the Local Group, UFDs, offer some of the best constraints on these processes. These environments can potentially be used to trace and constrain enrichment from the first stars \citep[e.g.][]{FrebelBromm2012,Ji2015,Ritter2015,Jeon2017,Hartwig2018}. and have been used to place constraints on the variety of possible astrophysical sources of r-process enrichment \citep[e.g.][]{Ji2016a,Ji2016b,Ji2018a,Ji2019,Tsujimoto2017,Duggan2018,Nagasawa2018,Ojima2018}. It is likely that chemical enrichment in more massive, metal rich systems with ongoing star formation like the Milky Way is much more uniform, averaging over the variety of complex hydrodynamics and feedback-driven turbulence interactions that may be driving much of the significant variance in stellar abundances seen at low metallicity. Stochasticity, which has been shown to be important for setting the width of stellar abundance patterns low metallicity environments of the Milky Way \citep[e.g.][]{Cescutti2008,Cescutti2014}, plays an even greater role in the evolution of these low mass dwarf galaxies \citep{Applebaum2018}. Indeed the increase in scatter in the stellar abundance patters of low mass dwarf galaxies and UFDs has been attributed to inhomogeneous mixing and stochastic effects \citep[e.g.][]{Norris2010a,Simon2015,Mashonkina2017,Suda2017}. This complicates the interpretation of abundance patters in these galaxies, particularly for constraining individual enrichment events.

% Maybe cite Norris2010a, Simon2015, Nissen1994 for abundance spreads in UFDs?

Recent cosmological zoom simulations have attained sufficient resolution to follow the evolution of individual UFDs (or their progenitors at high redshift) and can investigate their chemical properties \citep{Jeon2017,Corlies2018,Escala2018,Christensen2018}. Additional works have conducted more direct investigations into what drives the enrichment process for individual sources using both hydrodynamics simulations \citep{PanScannapiecoScalo2013,Ritter2015,Safarzadeh2017,Hirai2015,Hirai2017,Emerick2018b,HaynesKobayashi2019} and semi-analytic models \citep{Beniamini2018,KrumholzTing2018}. In \cite{Emerick2018b} we demonstrate that elements released through AGB winds have both wider abundance distributions in the ISM and lower ejection fractions from the galaxy than elements released in supernovae. It is becoming increasingly clear that enrichment is quite inhomogeneous in low mass dwarf galaxies (refs?), and also that the behavior of metals from any one enrichment event depends on a significant number of complex interactions. In spite of this progress, there is still substantial work to be done in understanding the physical processes that drive both evolution of both the mean and width of stellar abundances in low-mass dwarf galaxies.

@@

However, the metals from each source do gradually trend towards a fraction of $\sim$0.8 by the end of the 150~Myr simulation time, corresponding to the mass fraction of the CNM at this point in the galaxy's evolution (see Figure~\ref{ch1:figISM_evolution}). This trend -- that the fraction of metals contained in a given phases tends towards the mass fraction of that phase -- is true across all phases in the simulation. Therefore, we can consider that the metals for each source are well-mixed across the \textit{phases} of the ISM on timescales of $\sim$100-150~Myr. This is comparable, but less than, the dynamical timescale of the galaxy, $\sim$200~Myr. These two results suggest that the delay-time between when metals are available to enrich ongoing star formation is on order of $\sim$10~Myr across sources, and only significant for the lowest (E46, AGB-like) energy sources. Although well-mixed across phases, we emphasize that this does not imply that the abundances are the same across phases nor that the metals are spatially well-mixed across the galaxy. We investigate this second point further below.

@@

What this does suggest, however, is that in utilizing stars with unusual abundances at low metallicities to infer the yields of exotic enrichment sources, the spread of abundances in those stars may provide an important insight into the origin of those metals. For example, r-process enrichment from exotic enrichment events like hypernovae should manifest itself as more well-mixed than if the elements originated in lower energy NS-NS mergers. Conducting this analysis requires measured r-process abundances for a large sample of stars in an individual dwarf galaxy. This is challenging for the lowest mass, most metal poor dwarf galaxies, which typically only have a few to 10 stars with measured abundances, but could be done for more a few low mass dwarfs like Ursa Minor, and for more massive dwarfs like Carina, and Fornax \citep{Suda2017}.
% somerthing about needing cool onezone chemical models to try and model data with many free params might be nice here.

@@

Finally, examining the last panel of Figure~\ref{ch4:fig:SFR_comparison_CGM_CNM}, it does appear that the mixing efficiency of metals for each source increases with increasing global SFR. In the initial evolution, all events in run \runtwonu exhibit larger abundance spreads that take longer to begin mixing substantially than their counterparts in \runonenu. In general, \runtwonu~abundances remains larger for the first 100~Myr. The differences depend on the ejection energy, with the most significant difference found in comparing the E46 runs.
% May need more text here
